Viral Sequence Sorting Tools Evaluation

### 2022-09-15 TestingPipeline/Scripts/rule_tuning_2022.09.06.Rmd
Comparing high precision vs high recall pipeline types.

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +1

@@

```{r}
viruses$keep_score <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_kaiju = T,
                                              include_tuning = T,
                                              include_virsorter = T)
```

```{r}
viruses$confusion_matrix <- "true negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score<=1] <- "false negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score>=1] <- "true positive"
viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score>=1] <- "false positive"
```

@@

```{r}
p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_bar(stat="identity") +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()
p2
```
note: are these fungal actually false negatives? At least one was classified as viral by Kaiju.

>>>>

Virsorter2
    Viral >= 50: +0.5
    Viral >= 0.95: +0.5

@@

## high precision example
```{r}
viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_kaiju = T,
                                              include_tuning = T,
                                              include_virsorter = F)
```

```{r}
viruses$confusion_matrix_high_precision <- "true negative"
viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision<=1] <- "false negative"
viruses$confusion_matrix_high_precision[viruses$seqtype=="virus" & viruses$keep_score_high_precision>=1] <- "true positive"
viruses$confusion_matrix_high_precision[viruses$seqtype!="virus" & viruses$keep_score_high_precision>=1] <- "false positive"
```

visualizing confusion matrix by taxa
```{r}
confusion_by_taxa <- melt(table(viruses$confusion_matrix_high_precision, viruses$seqtype, viruses$Index))
colnames(confusion_by_taxa) <- c("confusion_matrix", "seqtype","Index", "count")
```

```{r}
ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_bar(stat="identity") +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()
```

```{r}
 ggplot(viruses, aes(x=checkv_viral_genes, y=confusion_matrix,
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_boxplot(alpha=0.3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Viral Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()

 ggplot(viruses, aes(x=percent_viral, y=confusion_matrix,
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_boxplot(alpha=0.3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Percent Genes Viral") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()

 ggplot(viruses, aes(x=hallmark, y=confusion_matrix,
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_boxplot(alpha=0.3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Hallmark Genes") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()

ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes,
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_point(alpha=0.3) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Hallmark Genes") +
  ylab("Number of Viral Genes") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()
```

```{r}
table(viruses$hallmark[viruses$confusion_matrix_high_precision=="false positive"]>0)

table(viruses$percent_host[viruses$confusion_matrix_high_precision=="false positive"]<50)
```

## high recall example
```{r}
viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_kaiju = T,
                                              include_tuning = T,
                                              include_virsorter = T)
```

```{r}
viruses$confusion_matrix <- "true negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_high_recall<=1] <- "false negative"
viruses$confusion_matrix[viruses$seqtype=="virus" & viruses$keep_score_high_recall>=1] <- "true positive"
viruses$confusion_matrix[viruses$seqtype!="virus" & viruses$keep_score_high_recall>=1] <- "false positive"
```

@@

```{r}
p2 <- ggplot(confusion_by_taxa, aes(x=count, y=as.factor(Index),
                   fill=confusion_matrix,
                   color=confusion_matrix)) +
  geom_bar(stat="identity") +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(4)), 0.5),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(4)), 1),
                    labels=c("false negative", "false positive", 
                             "true negative", "true positive")) +
  xlab("Number of Sequences") +
  ylab("") + 
  facet_wrap(~seqtype, scales = "free") +
  coord_flip()
p2
```
